R for Clinical Study Reports and Submission

### 2021-11-16 project-management.Rmd
Xiaoxia review project and submission Nov 15

Although many factors could decide if a project could execute efficiently,
we believe a few aspects are critical for long-term success,
especially when managing clinical data analysis projects at scale.

@@

For example, in the `esubdemo` project, we have 6 R Markdown files to create TLFs.

>>>>

Although many factors determine if a project can execute efficiently,
we believe a few aspects are critical for long-term success,
especially when managing clinical data analysis projects at scale.

@@

For example, in the `esubdemo` project, we have six R Markdown files to create TLFs.

### 2021-11-16 submission-package.Rmd
Xiaoxia review project and submission Nov 15

In this chapter, we will first give a high-level overview of what
assets in the eCTD submission package we should focus on when
submitting R code.
Secondly, we will discuss how to prepare the proprietary R packages
(if any), and make them part of the submission package.
Lastly, we will provide reusable templates for updating
Analysis Data Reviewer's Guide (ADRG) and Analysis Results Metadata (ARM)
so that the reviewers receive proper instructions to reproduce the
analysis results.

@@

The demo project (R package) we will prepare for submission is called
[`esubdemo`](https://github.com/elong0527/esubdemo), available on GitHub.
You can download or clone it:

@@

The demo submission package (not to be confused with the R package above)
is [`ectddemo`](https://github.com/elong0527/ectddemo), also available on GitHub.
You can download or clone it:

@@

We will assume the paths to the two folders are `esubdemo/` and `ectddemo/` below.

@@

A few things to pay attention to for passing compliance checks:

- The filenames under `programs/` should all use lower case
(with no underscores or other special characters).
- The `.txt` files should only contain ASCII characters.
This can be verified by `pkglite::verify_ascii()`
- Any `.docx` files should be converted to PDF files for formal submission.

Now you have a general idea about the relevant components of the
submission package.
We will start preparing the proprietary R packages in the following sections.

@@

Before we start, there are a few aspects to figure out,
to accurately identify the R packages for submission.

@@

Always use the same snapshot package repo for developing the TLFs and for submission.
Again, this can be checked in the project startup script,
discussed in section \@ref(reproduce).

@@

Introducing any extra external dependencies will likely increase the cost
of qualification, validation, and testing, and maintenance,
especially under Windows.
Therefore, it is recommended to keep the dependency chain simple,
especially when involving compiled code (e.g., C, C++, Fortran).

## Prepare R Packages for submission

To prepare R packages for submission, one needs to pack the packages into text files, and then verify if the file only contains ASCII characters.
With packed packages, one can unpack and install them from the text file, too.

@@

If successfully tested, this command will be used in the ADRG instructions
for restoring and installing the packed R package(s).

@@

After we converted the R packages and R Markdown files into the appropriate
formats and verified that they can be restored and executed correctly,
we should update the ADRG to provide guidelines on how to use them.

>>>>

In this chapter, we will first give a high-level overview of what
assets in the eCTD submission package we should focus on when
submitting R code.
Then, we will discuss how to prepare the proprietary R packages
(if any), and make them be part of the submission package.
In the end, we will provide reusable templates for updating
Analysis Data Reviewer's Guide (ADRG) and Analysis Results Metadata (ARM)
so that the reviewers receive proper instructions to reproduce the
analysis results.

@@

The demo project (R package) we will prepare for submission is called
[`esubdemo`](https://github.com/elong0527/esubdemo), which is available on GitHub.
You can download or clone it:

@@

The demo submission package (not to be confused with the R package above)
is [`ectddemo`](https://github.com/elong0527/ectddemo), which is also available on GitHub.
You can download or clone it:

@@

We assume the paths to the two folders are `esubdemo/` and `ectddemo/` below.

@@

A few things need to be paid attention to in order to pass compliance checks:

- The file names under `programs/` should be in lower case letters
(with no underscores or other special characters).
- The `.txt` files should only contain ASCII characters.
This can be verified by `pkglite::verify_ascii()`
- All `.docx` files should be converted to PDF files for formal submission.

Now you have a general idea about the relevant components of the
submission package.
We will prepare the proprietary R packages in the following sections.

@@

Before we start, there are a few aspects to figure out in order
to accurately identify the R packages for submission.

@@

Always use the same snapshot package repo for developing the TLFs and for submission.
Again, this can be checked in the project startup script,
as discussed in section \@ref(reproduce).

@@

Introducing any extra external dependencies will likely increase the cost
of qualification, validation, testing, and maintenance,
especially under Windows.
Therefore, it is recommended to keep the dependency chain simple,
especially when involving compiled code (e.g., C, C++, Fortran).

## Prepare R packages for submission

To prepare R packages for submission, one needs to pack the packages into text files, and then verify if the files only contain ASCII characters.
With packed packages, one can unpack and install them from the text files, too.

@@

If the test is successful, this command can be used in the ADRG instructions
for restoring and installing the packed R package(s).

@@

After we converted the R packages and R Markdown files into the appropriate
formats and verified that they can be restored and executed correctly,
we need update the ADRG to provide guidelines on how to use them.
